---
title: "2 Variable Analysis"
output: html_notebook
---

# In this problem set, you'll continue
# to explore the diamonds data set.

# Your first task is to create a
# scatterplot of price vs x.
# using the ggplot syntax.

# This assignment is not graded and
# will be marked as correct when you submit.

# ENTER YOUR CODE BELOW THIS LINE
# ===========================================


```{r}
data("diamonds")
```

```{r}
names(diamonds)
ggplot(aes(y = price, x = x), data = diamonds) +
  geom_point()
summary(diamonds$x)
?diamonds
```
I can see that some of them have length 0 but the have a price. They must be outliers or something going on with them.

and for the bulk of the data price increases with length of the diamond. 

majority of the data is in lengths 3 - 9

i see an exponential relationship between price and length.

```{r corelation between price, y, z and x}
cor.test(diamonds$price, diamonds$x)
cor.test(diamonds$price, diamonds$y)
cor.test(diamonds$price, diamonds$z)
```
# Create a simple scatter plot of price vs depth.

```{r}
ggplot(aes(y = price, x = depth), data = diamonds) +
  geom_point()
```

# Change the code to make the transparency of the
# points to be 1/100 of what they are now and mark
# the x-axis every 2 units. See the instructor notes
# for two hints.

```{r}
ggplot(aes(y = price, x = depth), data = diamonds) +
  geom_point(alpha = 1/100) + scale_x_continuous(breaks = seq(50,70,2))
```

```{r}
cor.test(diamonds$price, diamonds$depth)
```

The correlation coefficient is very less so, the two variables are not related to each other

# Create a scatterplot of price vs carat
# and omit the top 1% of price and carat
# values.

```{r}
ggplot(aes(x = carat, y = price), data = subset(diamonds, diamonds$price < quantile(diamonds$price, probs = 1 - 1/100))) + geom_point()
```

# Create a scatterplot of price vs. volume (x * y * z).
# This is a very rough approximation for a diamond's volume.

# Create a new variable for volume in the diamonds data frame.

```{r}
diamonds$volume <- (diamonds$x*diamonds$y*diamonds$z)
```


```{r}
ggplot(aes(x = volume, y = price), data = diamonds) +
  geom_point()
```
Most of the data has volume less than 1000.

and the volume seems to increase exponentially with price.

there is one extreme outlier in the data.

Some volumes are 0! There's an expensive diamond with a volume near 4000, and a cheaper diamond with a volume near 900

# exclude diamonds with volume 0 and >= 800
```{r}
with(subset(diamonds, diamonds$volume > 0 & diamonds$volume < 800), cor.test(price, volume))
new <- subset(diamonds, diamonds$volume > 0 & diamonds$volume < 800)
cor.test(new$price, new$volume)
```

# Adjust the transparency of the
# points and add a linear model to the plot. (See the
# Instructor Notes or look up the documentation of
# geom_smooth() for more details about smoothers.)

```{r}
ggplot(aes(x = volume, y = price), data = new) +
  geom_point(alpha = 1/20) + geom_smooth(method = 'lm', )
```

We can see a linear and exponential trend here till 400 there seems to be an exponential trend but the price staginates around 20000$ and we dont see any increase in price after that. so we cannot be entirely sure that this fits a linear model.


# Use the function dplyr package
# to create a new data frame containing
# info on diamonds by clarity.

# Name the data frame diamondsByClarity

# The data frame should contain the following
# variables in this order.

#       (1) mean_price
#       (2) median_price
#       (3) min_price
#       (4) max_price
#       (5) n

# where n is the number of diamonds in each
# level of clarity.

```{r}
library(dplyr)
diamondsByClarity <- diamonds %>%
  group_by(clarity) %>%
  summarise(mean_price = mean(price), median_price = median(as.numeric(price)), min_price = min(price), max_price = max(price), n = n()) %>%
  arrange(clarity)

```
# Your task is to write additional code to create two bar plots
# on one output image using the grid.arrange() function from the package
# gridExtra.

```{r}
diamonds_by_clarity <- group_by(diamonds, clarity)
diamonds_mp_by_clarity <- summarise(diamonds_by_clarity, mean_price = mean(price))

diamonds_by_color <- group_by(diamonds, color)
diamonds_mp_by_color <- summarise(diamonds_by_color, mean_price = mean(price))

```

```{r}
library(gridExtra)

p1 <- ggplot(aes(clarity, mean_price), data = diamonds_mp_by_clarity) + geom_col()
p2 <- ggplot(aes(color, mean_price), data = diamonds_mp_by_color) + geom_col()
grid.arrange(p1,p2,ncol = 2)
```
In the mean price by color the price increases with D > E < F < G < H < I < J

whereas in clarity there is a decreasing trend. with S12 having maximum mean price.

this is odd and against the trend that best color or clarity has more price.