---

title: "Problem_set_8"
author: "Sai Saketh Boyanapalli"
date: "October 6, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Welcome
Notes: Explore many vriables at once!

***

 Create a histogram of diamond prices.
 Facet the histogram by diamond color
 and use cut to color the histogram bars.
 
 Create a scatterplot of diamond price vs.
 table and color the points by the cut of
 the diamond.

### Scatterplot Review

```{r Scatterplot Review}
names(diamonds)
ggplot(aes(x = price), data = diamonds) +
  geom_histogram(aes(color = cut)) +
  facet_wrap(~ color, scales ="free_y" )
ggplot(aes(y = price, x = table), data = diamonds) +
  geom_point(aes(color = cut)) + xlim(50,70)
```

In the above plot we can see that for all colors prices tend to be skwed towards lower end.
and diamonds with color I and J tend to be more costlier. diamond cut is not very related to price. 
 Create a scatterplot of diamond price vs.
 volume (x * y * z) and color the points by
 the clarity of diamonds. Use scale on the y-axis
 to take the log10 of price. You should also
 omit the top 1% of diamond volumes from the plot.

 Note: Volume is a very rough approximation of
 a diamond's actual volume.
 Note: In the link, a color palette of type
 'div' was used to color the scatterplot using
 scale_color_brewer(type = 'div')

```{r}
diamonds$volume <- with(diamonds, x*y*z)
ggplot(aes(x = volume, y = price), data = subset(diamonds, volume < quantile(volume, probs = 99/100) )) +
  geom_point(aes(color = clarity)) + scale_y_log10()
```


Proportion of friendships initiated.

```{r}
pf$prop_initiated <- ifelse(pf$friend_count == 0,NA,pf$friendships_initiated/pf$friend_count)
```

 Create a line graph of the median proportion of
 friendships initiated ('prop_initiated') vs.
 tenure and color the line segment by
 year_joined.bucket.

 Recall, we created year_joined.bucket in Lesson 5
 by first creating year_joined from the variable tenure.
 Then, we used the cut function on year_joined to create
 four bins or cohorts of users.

 (2004, 2009]
 (2009, 2011]
 (2011, 2012]
 (2012, 2014]
 
 
```{r}
pf$year_joined <- floor(2014 - pf$tenure/365)
pf$year_joined_bucket <- cut(pf$year_joined, breaks = c(2004,2009,2011,2012,2014))
ggplot(aes(x = tenure, y = prop_initiated), data = pf) +
  geom_line(aes(color = year_joined_bucket), stat = 'summary', fun.y = median)
```
 Smooth the last plot you created of
 of prop_initiated vs tenure colored by
 year_joined.bucket. You can bin together ranges
 of tenure or add a smoother to the plot.

```{r}
ggplot(aes(x = tenure, y = prop_initiated), data = pf) +
  geom_smooth(aes(color = year_joined_bucket))
```
```{r}
with(subset(pf, year_joined > 2012), summary(prop_initiated, na.rm = T))
```
 Create a scatter plot of the price/carat ratio
 of diamonds. The variable x should be
 assigned to cut. The points should be colored
 by diamond color, and the plot should be
 faceted by clarity.
```{r}
ggplot(aes(x = cut, y = price/carat), data = diamonds) +
  facet_wrap(~clarity) +
  geom_point(aes(color = color)) +
  scale_color_brewer(type = 'div')
```


